---
title: "Darwin Core Mapping"
subtitle: "GBIF - A database of common vampire bat reports"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # for the date
library(inborutils)     # for utf 8
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv

```{r}

input_interim <- read_delim(here::here("datasets", "vampire-bat-occurrences", "data", "raw", "Desmodus_dataset_Dec_2021.csv" ), ",")

```

Preview data:

```{r}
input_interim %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_interim %<>% remove_empty("rows")
```

```{r}
input_interim %>% head(n = 5)
```

```{r}
input_interim <- inborutils::convertdf_enc(input_interim, from = "latin1", to = "UTF-8") 
```



# Occurrence core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim
```

## remove obsolete columns

## remove original GBIF data


```{r}
#occurrence %<>% filter(occurrence$datasetName == 'GBIF')

```

make link to GBIF records

```{r}
occurrence %<>% mutate(dwc_relatedREsourceID = if_else( datasetName == 'GBIF'
                                                    , str_c('https://www.gbif.org/occurrence/' , institutionID)
                                                    , '' ))

```

  
# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://doi.org/complete_this")
```

### add row numbers to genrate occurrenceID unique


```{r}
occurrence %<>% mutate(ID_row = row_number())
```


### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = institutionCode) # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```
```{r}
occurrence %<>% mutate(dwc_verbatimDatasetName = datasetName) 
```



### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "A database of common vampire bat reports") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = case_when (
                     datasetName == 'Lee'   ~ "MaterialCitation"
                    ,datasetName == 'Zarza' ~ "MaterialCitation"
                    ,basisOfRecord == 'UNKNOWN' ~ "occurrence"
                    ,datasetName == 'literature' ~ "MaterialCitation"
                    ,datasetName == 'Piaggio' ~ "MaterialCitation" 
                    ,datasetName == 'Juan Luis_Personal_Database' ~ "MaterialCitation"
                    ,datasetName == 'Literature' ~ "MaterialCitation"
                    ,datasetName == 'Bio_Diversi_Data_UY' ~ "MaterialCitation"
                    ,datasetName == 'PCMS' ~ "MaterialCitation"
                    ,datasetName == 'Literature' ~ "MaterialCitation"
                    ,datasetName == 'Streicker_Peru' ~ "MaterialCitation"  
                    ,datasetName == 'SAG' ~ "MaterialCitation" 
                    ,TRUE   ~ basisOfRecord ))
                
```


```{r}
occurrence %<>% mutate(dwc_scientificName = scientificName
                                )
```


```{r}
occurrence %<>% mutate(dwc_relatedREsourceID = if_else( datasetName == 'GBIF'
                                                    , str_c('https://www.gbif.org/occurrence/' , institutionID)
                                                    , '' ))

```


```{r}
##Occurrence %<>% mutate(basisOfRecord = replace(occurrence$basisOfRecord, 'UNKNOWN', 'HumanObservation'))
```


### informationWithHeld

### dataGeneralizations

### occurrenceID

```{r}
occurrence %<>% mutate(dwc_institutionCode = replace_na(institutionCode,'unk'))
```
```{r}
occurrence %<>% mutate(dwc_institutionID = replace_na(institutionID, 'unk'))
```


```{r}
vdigest <- Vectorize(digest)

occurrence %<>% mutate(dwc_occurrenceID = vdigest(str_c(ID_row, ":", dwc_institutionCode, ":" , dwc_institutionID), algo = "md5"))
  
```



### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = institutionCode)
```

### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount = individualCount)
```





### sex


```{r}
occurrence %<>%  mutate(dwc_sex = case_when(sex == 'M' ~'male'
                                          ,sex == 'F' ~'female'
                                          ,TRUE   ~ sex ))
                
                
```


### lifeStage


```{r}
occurrence %<>%  mutate(dwc_lifestage = lifeStage) 
                
```


### behavior

### occurrenceRemarks

```{r}
occurrence %<>% mutate(dwc_occurrenceRemarks = 'republished from https://doi.org/10.6084/m9.figshare.18745316')
```



### eventDate

eventDate already ok in source file (should be)

```{r}


#occurrence %<>% ymd(date)

occurrence %<>% mutate(dwc_eventDate = str_replace_all((paste( year,"-",month,"-",day))," ",""))
# mutate(eventDate = as_date(ymd(date))) # , format = "%Y-%m-%d")) 

head(occurrence, n = 5)
```

# some tests

# Data quality tests


parsed_names <- occurrence %>%
  distinct(scientificName)  %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function
speciesList <- parsed_names %>%
   select(scientificname)
write_csv(speciesList, here::here("datasets","memo-occurrences", "data", "processed", "specieslist.csv"), na = "")

```{r}
countries <- occurrence %>%
            distinct(country) %>%
            pull() 
view(countries)
```


# Location

```{r}
occurrence %<>%
  mutate(dwc_decimalLongitude = decimalLongitude) %>%
  mutate(dwc_decimalLatitude = decimalLatitude) %>%
  mutate(dwc_geodeticDatum = "WGS84") %>%
  mutate(dwc_coordinateUncertaintyInMeters = coorditeUncertaintyInMeters) %>%
  mutate(countryCode = case_when(  country == 'BRA' ~ 'BR'
                                  ,country == 'ARG' ~ 'AR'
                                  ,country == 'BLZ' ~ 'BZ'
                                  ,country == 'BOL' ~ 'BO'
                                  ,country == 'CHL' ~ 'CL'
                                  ,country == 'COL' ~ 'CO'
                                  ,country == 'CRI' ~ 'CR'
                                  ,country == 'ECU' ~ 'EC'
                                  ,country == 'GTM' ~ 'GT'
                                  ,country == 'GUF' ~ 'GF'
                                  ,country == 'GUY' ~ 'GY'
                                  ,country == 'HND' ~ 'HN'
                                  ,country == 'MEX' ~ 'MX'
                                  ,country == 'NIC' ~ 'NI'
                                  ,country == 'PAN' ~ 'PA'
                                  ,country == 'PER' ~ 'PE'
                                  ,country == 'PRY' ~ 'PY'
                                  ,country == 'SLV' ~ 'SV'
                                  ,country == 'SUR' ~ 'SR'
                                  ,country == 'TTO' ~ 'TT'
                                  ,country == 'URY' ~ 'UY'
                                  ,country == 'VEN' ~ 'VE'
                                  ,TRUE == coutry
                                  
                                  ))            
  
```

```{r}
head(occurrence, n = 5)
occurrence %<>%
  mutate(verbatimLongitude = round(verbatimLongitude)) %>%
  mutate(verbatimLatitude = round(verbatimLatitude)) %>%
  mutate(decimalLongitude = round(decimalLongitude, digits = 5)) %>%
  mutate(decimalLatitude = round(decimalLatitude, digits = 5))
```

```{r}
occurrence %<>%   
   mutate(decimalLatitude = as.character(format(decimalLatitude, nsmall = 5))) %>%
   mutate(decimalLongitude = as.character(format(decimalLongitude, nsmall = 5)))
```

### continent

```{r}
##occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
#occurrence %<>% mutate(countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality

municipality already in source file

```{r}
occurrence %<>%
  mutate(municipality = str_remove(municipality,"\\(BE\\)"))
```

### verbatimcoordinates

### verbatimLatitude

### verbatimLongitude

### verbatimcoordinatesystem

```{r}
#occurrence %<>% mutate(verbatimcoordinatesystem = "Lambert coordinates") # 
```

### verbatimSRS

```{r}
#occurrence %<>% mutate(verbatimSRS = "Belgian Datum 1972")
```

### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

### georeferenceRemarks

### identifiedBy

```{r}
occurrence %<>% mutate(identifiedBy = 'RATO')
```

### scientificName 

```{r}
occurrence %<>% mutate(scientificName = case_when(kind_nl == 'Muskusrat' ~ "Ondatra zibethicus",
                                   kind_nl == 'Bruine rat' ~ "Rattus norvegicus"
                                       )
                                )
```

### kingdom

```{r}
occurrence %<>% mutate(kingdom = "Animalia")
```

### scientificNameAuthorship

```{r}
occurrence %<>% mutate(scientificNameAuthorship = case_when(kind_nl == 'Muskusrat' ~ "Linnaeus, 1766",
                                   kind_nl == 'Bruine rat' ~ "Berkenhout, 1759"
                                       )
                                )


```

### taxonRank

```{r}
occurrence %<>% mutate(taxonRank = "species")
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(nomenclaturalCode = "ICZN") # e.g. "ICZN"
```

### occurrenceStatus

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(occurrenceStatus = case_when(action_nl == 'Vastgesteld' & individualCount > 0 ~ "Present",
                                        action_nl == 'Vastgesteld' & individualCount == 0 ~ "Absent",
                                        action_nl == 'Gevangen' & individualCount > 0 ~ "Present",
                                        action_nl == 'Gevangen' & individualCount == 0 ~ "Absent",
                                        action_nl == 'Verwijderd' & individualCount > 0 ~ "Present",
                                        action_nl == 'Verwijderd' & individualCount == 0 ~ "Absent",
                                        action_nl == 'Geen vangst' & individualCount > 0 ~ "Present",
                                        action_nl == 'Geen vangst' & individualCount == 0 ~ "Absent",
                                        action_nl == 'Sporen' & individualCount > 0 ~ "Present",
                                        action_nl == 'Sporen' & individualCount == 0 ~ "Absent",
                                        action_nl == 'Hoeveelheid'& individualCount > 0 ~ "Present",
                                        action_nl == 'Hoeveelheid'& individualCount == 0 ~ "Absent",
                                        individualCount > 0 ~ "Present",
                                        individualCount == 0 ~ "Absent",
                                        is.na(action_nl) | is.na(individualCount) ~ "Absent"
                                                  
                                       ))
head(occurrence, n = 5)
```

```{r eval=FALSE, include=FALSE}

occurrence %<>%
        filter(eventDate > '2018-01-12' & occurrenceStatus == "Present" & scientificName == "Ondatra zibethicus") 
        
head (occurrence, n = 5 )

```






## Post-processing

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
##occurrence %<>% select(-c(disposition, etiket_ref, collection, scientificName)) %<>% #remove collection columns
          ##  rename(scientificName = verbatimScientificName) 
```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

Define the order of the output columns

```{r}
col_order <- c( "type","language","license","rightsHolder","accessRights","datasetID"
               ,"institutionCode","datasetName","basisOfRecord","occurrenceID","recordedBy"
               ,"individualCount","organismQuantity","organismQuantityType","occurrenceStatus","occurrenceRemarks"
               ,"samplingProtocol","samplingEffort"
               ,"eventDate","continent","countryCode","municipality"
               ,"verbatimLatitude","verbatimLongitude","verbatimCoordinateSystem","verbatimSRS"
               ,"decimalLatitude","decimalLongitude","geodeticDatum","coordinateUncertaintyInMeters"
               ,"identifiedBy","scientificName","kingdom","scientificNameAuthorship","taxonRank","nomenclaturalCode"
               )
occurrence <- occurrence[, col_order]
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "mica-rato-occurrences", "data", "processed", "occurrence.csv"), na = "")
```
